
\renewcommand{\baselinestretch}{2}
\small\normalsize
\hbox{\ }
\vspace{-.65in}
\begin{center}
\large{Acknowledgments}
\end{center}
\vspace{1ex}

Throughout the time I have worked on this dissertation, I have had mentorship, help, and support, from many people. To all of them: Thank you very much. I am truly grateful.

I would first like to thank my advisor, Neil Spring. His mentorship has been invaluable in the making of this dissertation, and on a broader level, in the making of the researcher that I am today. His passion for research and attention to detail are qualities that I look up to and have learned to emulate. He has ever been a bastion of constructive criticism and his feedback has helped tremendously in honing my research skills. In addition, he has been flexible and accommodative of my occasionally unusual schedules. I could not have asked for more from an advisor.

There were several other mentors at the University of Maryland who had a direct role in shaping this dissertation. I have worked closely with Dave Levin and have benefitted much from his approach to research, his clear writing style, his brilliant presentations, his generosity, and his wonderful sense of humor. He has been a source of immense support and is an inspiration. Aaron Schulman helped me choose University of Maryland, introduced me to the problem I tackled in this dissertation, and importantly, also helped me sustain my research during difficult times. Aaron’s enthusiasm, curiosity, and work-ethic, inspired me to persist with graduate school. I am also very grateful to Bobby Bhattacharjee; his incisive questions and feedback throughout have helped this dissertation significantly and have also helped me grow as a researcher. Ashok Agrawala provided valuable guidance early on and Atif Memon gave helpful comments during the thesis proposal. I would also like to thank my committee members, John Dickerson and Mark Shayman, for their feedback on the dissertation manuscript and presentation. 

Mentors outside UMD have also provided valuable assistance during different stages of the dissertation. Alberto Dainotti has been instrumental during the finishing stages of the dissertation and has provided valuable feedback and support. Arthur Berger is another mentor who has helped considerably during the latter half of the dissertation; his thoughtfulness, ability to listen, and patience are qualities that I aim to emulate. I am also grateful to Matthew Luckie, Amogh Dhamdhere, and kc claffy, for introducing me to CAIDA and for helping me navigate the academic research world outside UMD. I would also like to thank Dave Plonka for showing me how to approach research with infectious enthusiasm. A special thank you to Krishna Sivalingam for giving me the opportunity to pursue my first research problem, and for his considerable help with graduate school applications. 

I am also very grateful to my colleagues and friends, who have made graduate school an enjoyable experience. Aaron, Randy, Lex, Yunus, Greg, Karla, and Jessy were brilliant colleagues who welcomed me to graduate school and helped me learn the ropes. Matt Lentz has been an extraordinary colleague and friend, helping me immensely with my research and presentations, while also being a participant in some of the more memorable conversations I’ve had. I am also very grateful to Zhihao Li for his friendship and support; I am especially privileged to have been his first co-author on a paper. I am also grateful to Philipp Richter for a delightful paper-writing experience together. James, Youndo, Stephen, Katura, Richie, thank you all for your support and encouragement, especially during the thesis proposal and dissertation defense. Thank you also for making the lab a great place to hang out. A special thank you to Brandi Adams for being a wonderful friend and for having shared in the journey. Sharron McElroy, thank you for the conversations and for making the reimbursement process enjoyable. My housemates and fellow PhD students, Bhaskar Ramasubramaniam, Amit Chavan, and Kartik Nayak: I am truly grateful for all the tea, food, help and support. Anshul Sawant, Manish Purohit, Meethu Malu, and Sudha Rao, thank you for all the memories over the years. And Kleoniki Vlachou, though your cakes were one of the highlights of my first couple of years of graduate school, it is your support and encouragement through the years that have been the true icing on the cake. 

I am also grateful to the students I’ve had the privilege to mentor and teach. Patrick Owen was the first undergraduate student I worked with and I am gratified that he is continuing to work on similar topics. Working with Ramakrishnan Sundara Raman and Reethika Ramesh was truly a delight; their enthusiasm and eagerness to learn made for an excellent mentoring experience. 

Last, but by no means the least, I would like to thank my family---including my grandparents, uncles, aunts, cousins, and even my nephews and nieces--- for their unceasing love and encouragement. I am especially grateful to my parents, Padmanabhan Raman and Lalitha Ramakrishnan for teaching me values like dedication, diligence, and optimism; without these values, I couldn’t have completed this dissertation. I am also immensely grateful to my parents for always giving me the freedom to pursue my dreams. Ranjani Padmanabhan, my cheerful and witty sister, has been ever ready with her support. My wife, Janani Saikumar, who has also been pursuing her PhD alongside me, is one of the principal reasons this dissertation was possible. From providing me with initial encouragement to pursue graduate school, to continuous love and support during the ebbs and flows of graduate life, she has played an immense part.
